\thetitle

### 2020-11-19 sun_phases/spin_qudit_tomo.tex
gave up on a tight reconstruction error bound due to difficulties with maximizing over the space of physical states with a non-negative spectrum

\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

@@

[\red{todo: comment on connection between $T_{\ell m}$ and $Y_{\ell m}$}]

@@

[\red{todo: add visualization of the signal recovery problem}]

@@

Since
\begin{enumerate*}
\item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
\item measuring spin projection along any single axis provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
\end{enumerate*}
one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
In the language of classical signal recovery, we would expect there to exist a fixed set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.  However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
Nonetheless, in the following section we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ [\red{todo: possibly refer to Section \ref{sec:axes} instead}].

@@

where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.

@@

where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.

@@

  \Bbk{\norm{\tilde\rho_V-\rho}^2}

@@

is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, such that $M_V M_V^{-1}=1$.
Note that the classical error scale $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time.
If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{error_scale/error_times.pdf}
  \caption{Serial runtime to compute the error scales $\S_V^2$ and $\epsilon_V^2$ with $\abs{V}=2d-1$ randomly chosen measurement axes.
    Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
    Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
    These results do not include fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of measurement axes $V$.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed using trace inner products of the {\it covariance matrices}
\begin{align}
  \C_\ell\sp{\rho} \equiv \sum_{m,n=-\ell}^\ell
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
\end{align}
where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it noise matrices}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
  \label{eq:error}
\end{align}
Maximizing this error over all $\rho$ (for a fixed set of axes $V$) yields a reconstruction error bound of the form
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ (defined below) is the same as that of computing the classical error scale $\S_V^2$ (see Figure \ref{fig:times}).
For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
We also note that the (squared) reconstruction error in Eq.~\eqref{eq:error} and its bound in Eq.~\eqref{eq:bound} obey the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

The quantum error scale $\epsilon_V^2$ only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.

The remainder of this section is devoted to defining the quantum error scale $\epsilon_V^2$ (derived in Appendix \ref{sec:error}), and can be skipped without consequence for the discussions in the rest of this paper.
To write out this error in full, we first rewrite the reconstruction error in Eq.~\eqref{eq:error} as
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
  \label{eq:error_simp}
\end{align}
where $\ket{\rho_\ell}\equiv\sum_m \rho_{\ell m}\ket{m}$, and similarly $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ is a vector of the scalars $\chi^V_{\ell m}$ defined by
\begin{align}
  \chi^V_{LM} &\equiv
  \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
  g_{L\ell} &\equiv \sum_{m,m'}
  \obk{T_{L,m+m'}|T_{\ell m}T_{\ell m'}} \op{m}{m'}, \\
  \I &\equiv \sum_m \p{-1}^m \op{-m}{m}, \\
  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
\end{align}
Here $g_{L\ell}$ is essentially a matrix of transition operator product coefficients (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.
Maximizing the reconstruction error in Eq.~\eqref{eq:error_simp} over all $\rho$ with $\tr\rho=1$ leads to the bound in Eq.~\eqref{eq:bound} with
\begin{align}
  \epsilon_V^2 \equiv \sum_{\ell>0} \sp{\f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
    + \f1d \tr\p{\N_{V\ell}}}.
\end{align}
Note that the maximum of Eq.~\eqref{eq:error_simp} may generally occur at an unphysical ``state'' $\rho$ with negative eigenvalues.
In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the restricted domain of states with a non-negative spectrum.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability [\red{todo: add figure}].
Two observations may explain the poor performance of these sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Revisiting the reconstruction error bound}
\label{sec:error}

>>>>

\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

@@

[\red{todo: investigate and comment on the connection between $T_{\ell m}$ and $Y_{\ell m}$?}]

@@

[\red{todo: add figure for signal recovery problem?}]

@@

Since
\begin{enumerate*}
\item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
\item measuring spin projection along any single axis provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
\end{enumerate*}
one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
In the language of classical signal recovery, we would expect there to exist a fixed set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.  However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
Nonetheless, in the following section we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ [\red{todo: refer to Section \ref{sec:axes} instead}].

@@

where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the transition operator $T_j\equiv T_{\v v\ell,0}$.

@@

where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials that estimate $\bk{T_j}_\rho$.

@@

  \E_V\p{\rho}^2
  \equiv \Bbk{\norm{\tilde\rho_V-\rho}^2}

@@

is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, satisfying $M_V M_V^{-1}=1$.
We refer to this error scale $\S_V^2$ as ``classical'' because the bound in Eq.~\eqref{eq:bound_eps} applies in the presence of classical sources of measurement error.
Note that the classical error scale $\S_V^2$ diverges if the measurement matrix $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$-sized blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time.
If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{error_scale/error_times.pdf}
  \caption{Serial runtime to compute the classical error scale $\S_V^2$ with $\abs{V}=2d-1$ randomly chosen measurement axes.
    Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
    Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
    These results do not include fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of measurement axes $V$.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:improved_bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the constraints that all $\bk{\Pi_{\v v\mu}}_\rho\ge0$ and $\tr\p\rho=1$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \E_V\p{\rho}^2 < \f{\epsilon_V^2}{n},
  &&
  \epsilon_V^2 \equiv \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:bound}
\end{align}
where the {\it quantum error scale} $\epsilon_V^2$ is defined in terms of (half of) the spectral range of $T_{\ell,0}$:
\begin{align}
  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}.
\end{align}
If $d$ is even, then $\Gamma_\ell=\max_\mu t_{\ell\mu}$.
For comparison with previous error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V^2$ is identical to that of $\S_V^2$.

Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
Though the individual bounds on $\bbk{\epsilon_{\v v\ell}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

The quantum error scale $\epsilon_V^2$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
  \label{eq:error}
\end{align}
where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$, and $\ket{\chi_{V\ell}}$ and $\N_{V\ell}$ are defined below.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.

The matrix $\N_{V\ell}$ that appears in Eq.~\eqref{eq:error} is defined in terms of the measurement matrix $M_{V\ell}$ and its right inverse $M_{V\ell}^{-1}$ by
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero.
The vector $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ in Eq.~\eqref{eq:error} is defined by
\begin{align}
  \chi^V_{LM} &\equiv
  \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
  g_{L\ell} &\equiv \sum_{m,m'}
  \obk{T_{L,m+m'}|T_{\ell m}T_{\ell m'}} \op{m}{m'}, \\
  \I &\equiv \sum_m \p{-1}^m \op{-m}{m}, \\
  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
\end{align}
Here $g_{L\ell}$ is a matrix of structure constants for the transition operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.

[\red{todo: comment on code availability}]

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:bound}.
If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$, defined in Eq.~\eqref{eq:reconstructed_state}.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
Finally, the error in $\tilde\rho_V$ can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho_V\to\tilde\rho_V$.

@@

As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability [\red{todo: add figure?}].
Two observations may explain the poor performance of these sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An improved reconstruction error bound}
\label{sec:improved_bound}
